---
title: "IoslidesDemo"
author: "Matthew DeLoia, Missional Analysis, ARI Program"
date: "8/13/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(Rmisc)
library(foreign)
library (memisc)
library(tidyverse)
library(readxl)
library(eRm)
library(ggcorrplot)
library(corrplot)
library(chron)
library(hms)
library(mclust)
library(imputeMissings)
library(randomForest)
library(plotly)
library(ggdendro)
library(cluster)
library(wordcloud)
library(wordcloud2)
library(stringr)
library(tm)
library(gganimate)
library(pracma)
library(lubridate)
library(plotly)

#slides available at: https://mattdeloia.github.io/DemoOUAug20/

likertNum <- function(x){
  case_when(
    x == "Very Accurate" ~ 5,
    x == "Moderately Accurate" ~ 4,
    x == "Neither Accurate Nor Inaccurate" ~ 3,
    x == "Moderately Inaccurate" ~ 2,
    x == "Very Inaccurate" ~ 1
  )
}

likertNum2 <- function(x){
  case_when(
    x == "VERY ACCURATE" ~ 5,
    x == "MODERATELY ACCURATE" ~ 4,
    x == "NEITHER INACC NOR ACCUR" ~ 3,
    x == "MODERATELY INACCURATE" ~ 2,
    x == "VERY INACCURATE" ~ 1
  )
}

#load Cyber Summit Questionnnaire data
df <- read_excel("DataScienceDemo_13Aug20.xls")
df <- df %>% rownames_to_column("ID") 
n_sample <- df %>% count()
n_sample  <- n_sample$n

#Load and Process Eugene-Springfield dataset
df_ESDemo <- as.data.set(spss.portable.file("demographics.por"), to.data.frame=T)
df_ESDemo <-as.data.frame(df_ESDemo)
df_ES <- as.data.set(spss.portable.file("IPIP2539-1.por"), to.data.frame=T)
df_ES <- as.data.frame(df_ES)
df_ES2 <- df_ES %>% dplyr::select (ID, A1:A99, C1:C9, M1:M9) %>% gather(A1:A99, C1:C9, M1:M9, key=Question, value = Response, na.rm = TRUE)
df_ES2$Response <- toupper(df_ES2$Response)
df_ES3 <- df_ES2 %>% 
  mutate_at(vars(`Response`), likertNum2) %>% 
  spread(key=Question, value=Response) 
df_ES4 <- df_ES %>%  dplyr::select(-(A1:A99),-(C1:C9), -(M1:M9)) %>% left_join(df_ES3, by="ID")
df_ES5 <- df_ESDemo %>% left_join(df_ES4, by="ID")
df_ES5$ID <- as.character(df_ES5$ID)
df_ESscored <- df_ES5 %>% group_by(ID) %>% 
  mutate(Anxiety = (sum(H926,H980, S3, H761, X107 ) ),      
         Social_Astuteness = (sum(H1137, P432, P451, V96, V153)),
         Breadth_of_Interest = (sum(X211, V140, V2, V330, H1265) ), 
         Complexity = (sum(H1276, Q228, H1267, H1284, V256) ), 
         Social_Confidence = (sum(H1110,S29,P364, H1366, H16 ) ), 
         Cooperativeness = (sum(P420,E147,D107,E68, H974)),  
         Empathy = (sum(D19, X253,H173,	D6,Q256) ), 
         Energy_Level = (sum(H54,	H555,H554,Q151,H79) ), 
         Innovation = (sum(H53,	V259,V226, H447,  X225) ),  
         Organization  = (sum(V98,Q163,V261, H258, V12) ), 
         Responsibility = (sum(H152,V303, D10, E80,V170) ), 
         Risk_Taking = (sum(S19,S4, R77, R58, Q126) ), 
         Sociability = (sum(D70,X161, H78, H985, Q238) ), 
         Tolerance = (sum(H148, Q59, X59, D113, X90) ),
         Traditional_Values = (sum(D88, V46, Q44, H296, X76) ) ) %>%
  dplyr::select (ID, SEX, AGE, EMPLOY, Anxiety:Traditional_Values) %>%
  dplyr::rename(Age = AGE, Gender = SEX) %>% na.omit() 
df_ESscored$Gender <- gsub("M", "Male", df_ESscored$Gender )
df_ESscored$Gender <- gsub("F", "Female", df_ESscored$Gender )

#filter to desired comparative population
df_ESscored2 <- df_ESscored %>% filter(EMPLOY %in% c("full time", "part time", "retired"), Age>=22 & Age<=75) %>%  dplyr::select(ID, Gender, Age, Anxiety:Traditional_Values )  %>% na.omit()
summaryES <- df_ESscored2 %>% gather(Anxiety:Traditional_Values, key=Dimension, value=Score) %>% summarySE(groupvars = "Dimension", measurevar = "Score" )

anxiety <- summaryES %>% filter(Dimension=="Anxiety")
breadth_of_interest <- summaryES %>% filter(Dimension=="Breadth_of_Interest")
complexity <- summaryES %>% filter(Dimension=="Complexity")
cooperativeness <- summaryES %>% filter(Dimension=="Cooperativeness")
empathy <- summaryES %>% filter(Dimension=="Empathy")
energy_level <- summaryES %>% filter(Dimension=="Energy_Level")
innovation <- summaryES %>% filter(Dimension=="Innovation")
organization <- summaryES %>% filter(Dimension=="Organization")
responsibility <- summaryES %>% filter(Dimension=="Responsibility")
risk_taking <- summaryES %>% filter(Dimension=="Risk_Taking")
sociability <- summaryES %>% filter(Dimension=="Sociability")
social_astuteness <- summaryES %>% filter(Dimension=="Social_Astuteness")
social_confidence <- summaryES %>% filter(Dimension=="Social_Confidence")
tolerance <- summaryES %>% filter(Dimension=="Tolerance")
traditional_values <- summaryES %>% filter(Dimension=="Traditional_Values")
responsibility <- summaryES %>% filter(Dimension=="Responsibility")

#Processing Demograpic data (Age)
medianAge <- df %>% select(ID, Age) 
medianAge <- median(medianAge$Age)

#Processing Demograpic data (Education)
df_ed <-df %>% select(Education) %>%   group_by(Education) %>% summarise(n=n()/n_sample)
df_ed$Education <-  factor(df_ed$Education, levels=c("Secondary Education", "Associates", "Bachelors", "Masters", "PhD"))

#Processing Demograpic data (Hobbies)
df_hobby <-df %>% select(Hobbies) 
df_hobby$Hobbies<- gsub("\\(", "", df_hobby$Hobbies)
df_hobby$Hobbies<- gsub("\\)", "", df_hobby$Hobbies)
df_hobby2<-df_hobby %>% separate(Hobbies, c("Hobbies1", "Hobbies2", "Hobbies3", "Hobbies4", "Hobbies5", "Hobbies6", "Hobbies7", "Hobbies8"), sep=",") %>%
  gather(Hobbies1:Hobbies8, key=Category, value=Hobbies) %>% select (-Category ) %>%
  na.omit(df_hobby) 




#Scoring respondant data        
df_scores <- df %>%   
  mutate_at(vars(P_AN5:P_RE4), likertNum) 


#Summary scores for personality and cognitive areas
df_scored <- df_scores %>% dplyr::group_by(ID) %>% 
  mutate(Anxiety = (sum(P_AN1, P_AN2, P_AN3, P_AN4, P_AN5 ) ) ,
         Social_Astuteness = (sum(P_AS1, P_AS2, P_AS3, P_AS4, P_AS5	)),
         Breadth_of_Interest = (sum(P_BR1, P_BR2, P_BR3, P_BR4, P_BR5	) ) , 
         Complexity = (sum(P_CM1, P_CM2, P_CM3, P_CM4, P_CM5) ) ,
         Social_Confidence = (sum(P_CD1, P_CD2, P_CD3, P_CD4, P_CD5) ) ,
         Cooperativeness = (sum(P_CO1, P_CO2, P_CO3, P_CO4, P_CO5)),
         Empathy = (sum(P_EM1, P_EM2, P_EM3, P_EM4, P_EM5) ) , 
         Energy_Level = (sum(P_EN1, P_EN2, P_EN3, P_EN4, P_EN5)) ,
         Innovation = (sum(P_IN1, P_IN2, P_IN3, P_IN4, P_IN5) ) , 
         Organization = (sum(P_OR1, P_OR2, P_OR3, P_OR4, P_OR5 ) ) , 
         Responsibility = (sum(P_RE1, P_RE2, P_RE3, P_RE4, P_RE5) ) ,
         Risk_Taking = (sum(P_RI1, P_RI2, P_RI3, P_RI4, P_RI5	) ) , 
         Sociability = (sum(P_SO1, P_SO2, P_SO3, P_SO4, P_SO5 ) ) , 
         Tolerance = (sum(P_TO1, P_TO2, P_TO3, P_TO4, P_TO5 ) ) ,
         Traditional_Values= (sum(P_TR1, P_TR2, P_TR3, P_TR4, P_TR5) ))  %>%  
  select(ID, Name, Anxiety:Traditional_Values, Email) 

#Scale socres to ES sample (n=151)
df_scored2 <- df_scored %>%
  group_by(ID) %>%
  mutate(Anxiety= (Anxiety-anxiety$Score) / anxiety$sd)   %>% 
  mutate(Breadth_of_Interest= (Breadth_of_Interest-breadth_of_interest$Score)/breadth_of_interest$sd)  %>% 
  mutate(Complexity= (Complexity - complexity$Score)/ complexity$sd)  %>% 
  mutate(Cooperativeness= (Cooperativeness-cooperativeness$Score)/cooperativeness$sd)  %>% 
  mutate(Empathy= (Empathy-empathy$Score)/empathy$sd) %>% 
  mutate(Energy_Level= (Energy_Level-energy_level$Score)/energy_level$sd) %>% 
  mutate(Innovation= (Innovation-innovation$Score)/innovation$sd)  %>%
  mutate(Organization= (Organization-organization$Score)/organization$sd) %>% 
  mutate(Responsibility= (Responsibility-responsibility$Score)/responsibility$sd) %>% 
  mutate(Risk_Taking= (Risk_Taking-risk_taking$Score)/risk_taking$sd)  %>% 
  mutate(Sociability= (Sociability-sociability$Score)/sociability$sd) %>%
  mutate(Social_Astuteness= (Social_Astuteness-social_astuteness$Score)/social_astuteness$sd) %>%
  mutate(Social_Confidence= (Social_Confidence-social_confidence$Score)/social_confidence$sd)  %>%
  mutate(Tolerance= (Tolerance-tolerance$Score)/tolerance$sd) %>%
  mutate(Traditional_Values= (Traditional_Values-traditional_values$Score)/traditional_values$sd)

df_scored2 <- as.data.frame(df_scored2)
```

## Introduction {.build}
Problem Statement:  To understand personality differences between technologists and the general workforce.

Hypothesis to test: Mission Analysis and Enablement workers are similar in personality to the general workforce population.

## Background (1 of 2)| Personality Assessment

This 75 question personality assessment aims to measure 15 underlying personality traits using similar measurements to the <a href="https://w8y7a2j5.stackpathcdn.com/wp-content/uploads/2015/03/JPI-R-Scales.pdf">Jackson Personality Inventory - Revised (JPI-R)</a>.  

The JPI-R measures were selected due to the range of personality traits measured and inclusion of 7 measurements for "analytical" and "dependable," traits previously reported in those performing highly technical work roles.  This set of 15 measures has been useful in vocational counseling.  

This ***JPI-R proxy*** consists of questions drawn from the public domain repository of personality questions  (https://ipip.ori.org/). <br>

## Background (2 of 2)
<div class="columns-2">
***Analytical***

- Complexity 
- Breadth of Interest
- Innovation
- Tolerance

***Extroverted***

- Sociability 
- Social Confidence 
- Energy Level 

***Emotional*** 

- Empathy 
- Anxiety 
- Cooperativeness 

***Opportunistic*** 

- Risk Taking 
- Social Astuteness 

***Dependable***

- Organization 
- Traditional Values 
- Responsibility
</div>

<div class="notes">

- Complexity: seeks intricate solutions to problems
- Tolerance: accepts people even though they may differ from oneself; welcomes dissent
- Cooperativeness: is susceptible to social influence and group pressures; follows suit
- Social Astuteness: skillful at persuading others to achieve a particlular goal; occasionally seen as manipulative of others
- Traditional Values: values traditional customs and beliefs; others may view individual as "old-fashioned"
- Responsibility: strong obligation to be honest and upright; an inflexibly conscience

</div>

## Education (n=`r n_sample`)

```{r Demographics, echo=FALSE, message=FALSE}
df_ed %>% ggplot(aes(x=Education, y=n)) + 
  geom_col(fill="skyblue") + 
  ylab("Percentage") + 
  geom_text(aes( x=Education, y=n, label = round(n, 2), ),color="black", vjust= -.1, hjust = .3, size=3.5) +
  theme(axis.text.y=element_blank(), axis.ticks.y=element_blank()) + ylim(0,1) + xlab("")
```

## Age
Median age is `r medianAge` years.
```{r Demographics2, echo=FALSE, message=FALSE}
df %>% ggplot(aes(x=Age)) +
  geom_density(fill="skyblue") + 
  theme(axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  ylab("") + xlab ("") 
```

## Hobbies 
```{r Hobbies, message=FALSE, warning=FALSE, echo=FALSE}
text2 <- df_hobby2$Hobbies# Create a corpus  
docs2 <- Corpus(VectorSource(text2))
docs2 <- docs2 %>%
  #tm_map(removeNumbers) %>%
  tm_map(removePunctuation) %>%
  tm_map(stripWhitespace)
docs2 <- tm_map(docs2, content_transformer(tolower))
docs2 <- tm_map(docs2, removeWords, stopwords("english"))
dtm2 <- TermDocumentMatrix(docs2) 
matrix2 <- as.matrix(dtm2) 
words2 <- sort(rowSums(matrix2),decreasing=TRUE) 
df_text2 <- data.frame(word = names(words2),freq=words2)
wordcloud2(data=df_text2, size=.4, color='random-dark', backgroundColor = "lightgray" )
```

## Personality Results 
```{r PersonalityResults, echo=FALSE, message=FALSE}
personality_results <- df_scored2  %>% 
  gather(Anxiety:Traditional_Values, key=Dimension, value = Score) %>%
  summarySE(measurevar = "Score", groupvars = "Dimension") %>%
   mutate(Comparison = if_else(Score<=-.5, "Low", if_else(Score>=.5, "High", "Average")))
personality_results$Comparison <- factor (personality_results$Comparison, levels = c("High", "Average", "Low"))

personality_results %>% 
        mutate(Dimension=reorder(Dimension, Score, FUN=max)) %>%    
        ggplot(aes(x=Dimension, y=round(Score, 2), color=Comparison)) + 
        geom_point() +
        geom_errorbar(aes(ymin=Score-ci, ymax=Score+ci), width=.1 ) +
        coord_flip() + ylim(-2,2) +
        geom_hline(yintercept=0, linetype="dotted") +
        xlab("") +
     ylab("group mean with 95% ci (z score)") +
        scale_color_manual(name="general workforce comparison", values=c("green", "darkgray", "blue")) +
  theme(legend.title= element_text(color="black", size=10), legend.position = "top") + 
  labs(caption = "Note: z scores developed from the E-S Community Sample data to establish norms of the general workforce")
```
<div class="notes">

What aspects of personlity are most stable or like the general workforce?

</div>

## Personality Results (2 of 2)
```{r PersonalityResults2, echo=FALSE, message=FALSE}
p <- df_scored2 %>% gather(Anxiety:Traditional_Values, key=Dimension, value=Score) %>%  ggplot(aes(x=reorder(Dimension, Score, .fun=mean), y=Score)) + 
  geom_point(size=4,  color="skyblue") + 
  coord_flip() + ylim(-3,3) +
  geom_hline(yintercept = 0, linetype="dashed") +
  xlab("") + ylab("z score") + 
  theme(legend.position = "top") 

anim <- p + transition_states(ID, transition_length = .5, state_length = 1) 
animate(anim, height=500, width=600)
```

## Clustering Analysis

```{r HClustering, echo=FALSE, warning=FALSE}
clustering <- df_scores  %>% select(P_AN5:P_RE4, Name) %>% column_to_rownames ("Name")
h_cluster <- hclust(dist(clustering))
h_cluster %>% ggdendrogram(rotate=TRUE, theme_dendro = TRUE, size=1)
```

## Conclusions {.build}
<div class="blue">
Hypothesis to test: Mission Analysis and Enablement workers are similar in personality to the general workforce population.
</div>


What is your Conclusion: ***Accept*** the Null Hypothesis -or- ***Reject*** the Null Hypothesis?

## References
Douglas N. Jackson, Technical Manual: Jackson Personlaity Inventory - Revised, published by Sigma Assessment Systems, August 1994.

International Personality Item Pool, available at: https://ipip.ori.org/

R Core Team (2019). R: A language and environment for statistical
  computing. R Foundation for Statistical Computing, Vienna, Austria. URL
  https://www.R-project.org/.

Eugene-Springfield Community Sample, Harvard Dataverse, available at: https://dataverse.harvard.edu/dataverse/ESCS-Data 

Full Presentation available at: https://mattdeloia.github.io/DemoOUAug20/ 

## Back-up: Gender Differences
```{r TestComparison, echo=FALSE, message=FALSE, warning=FALSE}
df_ESscored <- as.data.frame(df_ESscored)
df_ESscored %>% mutate_at(vars(Anxiety:Traditional_Values), scale) %>%  gather(Anxiety:Traditional_Values, key=Dimension, value=Score) %>% summarySE(groupvars = c("Gender", "Dimension"), measurevar = "Score") %>% 
      mutate(Dimension=reorder(Dimension, Score, FUN=max)) %>%
      ggplot(aes(x=Dimension, y=Score, group=Gender)) +
      geom_errorbar(aes(ymin=Score-ci, ymax=Score+ci), width=.2 ) +
      geom_point( aes(color=Gender), size=3) +
      coord_flip() + ylim(-2,2) +
      scale_color_manual(values=c("tomato", "lightblue")) +
      ylab("gender mean with 95% ci (z score)") +
            xlab("") +  
      theme(legend.position = "top") + 
      labs(caption = "*gender differences by personality trait based on E-S Community Sample data (n~400)")
```


## Back-Up: Trait Descriptions (1 of 3) 
Analytical <br>

 - Complexity: seeks intricate solutions to problems
 - Breadth of Interest: interested in learing about a diversity of things
 - Innovation: motivated to develop novel solutions to problems; values new ideas
 - Tolerance: accepts people even though they may differ from oneself; welcomes dissent
 
## Back-Up: Trait Descriptions (2 of 3) 
Emotional 

 - Empathy: tends to identify closely with other people and their problems
 - Anxiety: tends to worry over inconsequential matters; apprehensive about the future
 - Cooperativeness: is susceptible to social influence and group pressures; follows suit

Extroversion 

- Sociability: eagerly join a varity of social groups; actively social
- Social Confidence: confident in dealing with others; shows presence in interpersonal situations
- Energy level: is active and spirited; does not tire easily; capable of intense work for long periods

## Back-Up: Trait Descriptions (3 of 3) 
Opportunistic 

- Social Astuteness: skillful at persuading others to achieve a particlular goal; occasionally seen as manipulative of others, but ordinarily diplomatic
- Risk Taking: enjoys gambling and taking a chance; willingly exposes self to situations with uncertain outcomes

Dependable 

- Organization: makes effective use of time; not easily distracted
- Traditional Values: values traditional customs and beliefs; others may view individual as "old-fashioned"
- Responsibility: strong obligation to be honest and upright; an inflexibly conscience   
